MT @HanBaldwin: Fightin OOVs in German \#twitter

### 2013-08-12 darmstadt_gscl_2013/article_twitter_normalization.tex
statistics for spell class in hunspell updated

In the next Section we will give a short overview of existing
scientific approaches to the problem of tackling out-of-vocabulary
words in non-standard texts. After that, in Section
\ref{error:analysis} we will analyze which types of noisiness
phenomena are especially characteristic for German Twitter. Section
\ref{normalization} will subsequently describe an automatic procedure
for mitigating some of the most prominent of those effects. In a
concluding step, we will perform an evaluation of the results of this
procedure and give further suggestions for future research.

@@

The reason for this confusion is the fact that the above
classifications both rely on several independent criteria at the same
time, though each of these criteria characterizes an NTN system from a
different point of view. As a consequence, unambiguous assignment of
an NTN approach to one particular class often becomes impossible. In
order to avoid this, we instead suggest using separate classifications
for each type of involved criteria which might characterize an NTN
technique.

@@

Here once again the question of classification arose. This time with
regard to the reasons why different tokens could have been omitted
from corresponding applications' dictionaries. In this respect,
division into following groups seemed to us to be appropriate to us:
\begin{enumerate}
  \item \textbf{Objective limitedness of machine-readable dictionary
    (MRD)}. Among this group, we counted words of basic vocabulary
    which did not get into applications' MRD either because they
    supposedly were rare or because they did not exist at the time
    when dictionaries were created. Another reason for inclusion in
    this type was the belonging of a word to an open lexical or
    part-of-speech class (like, for example, named entities or
    interjections) which are often omitted from MRDs due to the
    impossibility to fully cover them;\label{dict}
  \item \textbf{Sloppiness of users' input}. In the scope of this
    second group, we considered incorrect spellings of words
    encountered in text;\label{spell}
  \item \textbf{Stylistic specifics of text genre}. This group
    comprised words which could be considered as illegal from the
    point of view of standard language but were perfectly valid terms
    in the domain of web discourse or more specifically in Twitter
    communication.\label{style}
\end{enumerate}

@@

We subdivided class \ref{dict} into the following subclasses:
\begin{enumerate}
\item regular German words, e.g. \textit{aufm},
  \textit{losziehen};\label{regular}
\item compounds, e.g.  \textit{Altwein},
  \textit{Amtsapothekerin};\label{compound}
\item abbreviations, e.g. \textit{NBG}, \textit{OL};\label{abbr}
\item interjections, e.g.  \textit{aja}, \textit{haha};\label{inj}
\item named entities, with subclasses:\label{ne}
  \begin{enumerate}
  \item persons, e.g.  \textit{Ahmadinedschad}, \textit{Schweiger};
  \item geographic locations, e.g.  \textit{Biel}, \textit{Limmat};
  \item companies, e.g. \textit{Apple}, \textit{Facebook};
  \item product names, e.g. \textit{iPhone}, \textit{MacBook};
  \end{enumerate}
\item neologisms, with subclasses:\label{neolog}
  \begin{enumerate}
    \item newly coined German terms, e.g. \textit{entfolgen},
      \textit{gegoogelt};\label{new}
    \item loanwords, e.g. \textit{Community},
      \textit{Stream};\label{loan}
  \end{enumerate}
\item and, finally, foreign words like \textit{is} or \textit{now}
  which in contrast to \ref{loan} were not mentioned in any existing
  German lexica and did not comply with inflectional rules of German
  grammar.\label{fw}
\end{enumerate}
Though this division is admittedly arbitrary to a certain degree and
also has the disadvantage of simultaneously involving different
linguistic criteria, the underlying notion here was simple -- valid
words could have been omitted from an MRD either due to the
limitations of developers' capacities (group \ref{regular}), active
word formation processes or lexical productivity of the language
itself (groups \ref{compound} through \ref{new}) or also due to
language's openness to foreign language systems (groups \ref{loan} and
\ref{fw}).

@@

As is clear from the table, deletions are by far the most common type
of misspellings. The reasons for that are either relatively frequent
omissions of characters made by users or even more often automatic
truncations of too long messages which are performed by Twitter
service itself.\footnote{As is generally known, Twitter imposes a
  strong restriction on the length of posted messages which can be no
  longer than 140 characters. Upon exceeding this length, tweets get
  automatically truncated to the maximal allowed length.}

Finally, the third group -- ``Stylistic specifics of text genre'' --
was subdivided into the subgroups:
\begin{enumerate}
  \item @-tokens, e.g. \textit{@ZDFonline};
  \item hashtags, e.g. \textit{\#Kleinanzeigen};
  \item links, e.g. \textit{http://t.co};
  \item smileys, e.g. \textit{:-P};
  \item slang, e.g. \textit{OMG}.
\end{enumerate}
according to the formal or lexical class which the tokens belonged
to. Additionally, we also considered as \textit{slang} spelling
variants of standard language words which could be regarded as their
colloquial equivalents. Such words included cases like, for example,
\textit{ned} instead of \textit{nicht} or \textit{grad} instead of
\textit{gerade} and were marked both as \textit{misspelling}s and
\textit{slang} in our classification. A detailed statistics on the
subgroups of class 3 is shown in Table \ref{table:style}:
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``Stylistic specifics of text genre''\label{table:style}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    @-tokens & 13.02 & 20.23 & 16.19 & 21.91\\
    hashtags & 7.35 & 6.18 & 13.06 & 10.59\\
    links & 2.43 & 0.4 & 4.89 & 6.07\\
    smileys & 2 & 0.73 & 6.88 & 1.2\\
    slang & 16.22 & 5.27 & 6.94 & 4.77\\\hline
    {\bfseries total} & 41.02 & 32.81 & 47.96 & 44.54\\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

A striking outlier of 16.22~\% for slang tokens in column 1 of the
Table is explained by the fact that the word ``RT'' which occurred
1,235 times in our texts and was by far the most frequent OOV in
analyzed data, was recognized as out-of-vocabulary by
\texttt{hunspell} but was not deemed as such by
\texttt{TreeTagger}. Luckily, such singleton cases were rather rare
exceptions during the analyses and did not affect much the remaining
classes of OOVs, so that distributions of unknown tokens for both
tools were more or less similar or at least comparable to a certain
extent.

But an even more important conclusion which can be drawn from the
analyzed data is the fact that for both \texttt{TreeTagger} and
\texttt{hunspell}, Twitter-specific phenomena like special tokens,
colloquial expressions or sloppily typed words accounted for more than
a half of all unknown words found during the analysis. Since different
classes of these Twitter phenomena were formed by different processes
and also have different degrees of ambiguity, we will suggest
different normalization procedures for each of them in the next
Section.

@@

\subsection{Replacement of Twitter-Specific Phenomena}
In August 2007, Chris Messina, a renowned advocate of open-source
community, sent out a Twitter message suggesting that the pound sign
could be used for organizing groups on Twitter (cf. Parker,
\cite{parker}). The ``hash godfather'', as Chris Messina described
himself, allegedly borrowed this idea from Internet Relay Chats (IRC)
-- the forebears of modern social networks -- which have been using
the ``\#{}''-sign to mark channel names since the early 1990s.

Luckily for us, this ``novelty'' by Messina as well as @-tokens which
are used in Twitter to mark names of user account, both have strict
formal features by which these tokens could be identified. Other types
of words which are usually considered as OOVs and also have
unambiguous formal traits are hyperlinks, e-mail addresses, and
smileys. The presence of such formal criteria suggests that these
classes could best be handled by rule-methods and namely finite-state
techniques like finite-state transducers (FST) (cf. Jurafsky and
Martin, \cite{jurafsky}, pp.57-60).

For our purposes, we developed a prototypic Python system analogous to
FSTs in which a set of regular expressions was associated with
corresponding actions performed on matched subgroups. In this system,
absolutely unambiguous tokens like web links and e-mail addresses were
replaced with special words ``\%Mail'' and ``\%Link''
correspondingly. Emoticons were replaced with tokens ``\%PosSmiley'',
``\%NegSmiley'', depending on the type of emotion presumably conveyed
by these expressions. In cases when an emoticon was ambiguous with
regard to its polarity, it was replaced with the artificial word
``\%Smiley''.

Additionally to that, problematic UTF-8 characters like special quotes
and ellipsis which caused incorrect work of tokenizer were converted
to their corresponding ISO-8859-1 equivalents.

For hashtags, we simply stripped off the pound character from the
beginning of such words thus leaving only the alphabetic part of
tokens. A more complicated case turned out to be the @-tokens. For
them, we had to decide depending on the context whether these tokens
played an important syntactic role in sentence or not. In cases they
did, we additionally looked whether the alphabetic part following the
``@'' character was present in dictionary. If it was, we only stripped
the ``at'' character off, otherwise whole word was replaced with token
``\%UserName''. In cases when @-tokens served as addresses or retweet
marks, we deleted them from message completely since they were of
little interest and importance for further syntactic and semantic
analysis and could hardly be assigned a proper part-of-speech tag or
syntactic role.

All the artificially introduced words were added to custom user
dictionaries of \texttt{TreeTagger} and \texttt{hunspell}. For
tagging, we assigned NE tags to all artificial words except for
emoticon replacements which got the tag INJ. Additionally, positions
and lengths of all replacements we remembered along with original
words which were replaced and a restoration step to original forms was
performed for certain classes of tokens after tagging.

@@

%
\subsection{Restoration of misspellings}
%
\subsubsection{Restoration of Umlauts}
%
\subsubsection{Squeezing of Elongated Words}
%
\subsubsection{Translation of Slang Misspellings}
%
\section{Evaluation}\label{evaluation}

>>>>

In the next Section, we will give a short overview of existing
scientific approaches to the problem of tackling out-of-vocabulary
words in non-standard texts. After that, in Section
\ref{error:analysis} we will analyze which types of noisiness
phenomena are especially characteristic for German Twitter. Section
\ref{normalization} will subsequently describe an automatic procedure
for mitigating some of the most prominent of those effects. In a
concluding step, we will perform an evaluation of the results of this
procedure and give further suggestions for future research.

@@

This confusion is explained by the fact that the above classifications
both rely on several independent criteria at the same time, however
each of these criteria characterizes an NTN system from a different
point of view. As a consequence of this, unambiguous assignment of an
NTN approach to one particular class often becomes impossible. In
order to avoid this, we instead suggest using separate classifications
for each type of involved characterizing criterion.

@@

We classified found OOV tokens into the following three groups
according to the reasons why these tokens could have been omitted from
corresponding applications' dictionaries:
\begin{enumerate}
  \item \textbf{Objective limitedness of machine-readable dictionaries
    (MRD)}. Among this group, we counted words of basic vocabulary
    which did not get into applications' MRD either because they
    supposedly were rare or did not exist at the time when
    dictionaries were created. Another reason for inclusion in this
    type was the belonging of a word to an open lexical or
    part-of-speech class (like, for example, named entities or
    interjections) which are often omitted from MRDs due to the
    impossibility to fully cover them;\label{dict}
  \item \textbf{Sloppiness of users' input}. In the scope of this
    second group, we considered incorrect spellings of words
    encountered in text;\label{spell}
  \item \textbf{Stylistic specifics of text genre}. This group
    comprised words which could be considered as illegal from the
    point of view of standard language but were perfectly valid terms
    in the domain of web discourse or more specifically in Twitter
    communication.\label{style}
\end{enumerate}

@@

We subdivided class \ref{dict} into the following subclasses:
\begin{enumerate}
\item regular German words, e.g. \textit{meins},
  \textit{losziehen};\label{regular}
\item compounds, e.g.  \textit{Altwein},
  \textit{Amtsapothekerin};\label{compound}
\item abbreviations, e.g. \textit{NBG}, \textit{OL};\label{abbr}
\item interjections, e.g.  \textit{aja}, \textit{haha};\label{inj}
\item named entities, with subclasses:\label{ne}
  \begin{enumerate}
  \item persons, e.g.  \textit{Ahmadinedschad}, \textit{Schweiger};
  \item geographic locations, e.g.  \textit{Biel}, \textit{Limmat};
  \item companies, e.g. \textit{Apple}, \textit{Facebook};
  \item product names, e.g. \textit{iPhone}, \textit{MacBook};
  \end{enumerate}
\item neologisms, with subclasses:\label{neolog}
  \begin{enumerate}
    \item newly coined German terms, e.g. \textit{entfolgen},
      \textit{gegoogelt};\label{new}
    \item loanwords, e.g. \textit{Community},
      \textit{Stream};\label{loan}
  \end{enumerate}
\item and, finally, foreign words like \textit{is} or \textit{now}
  which in contrast to \ref{loan} were not mentioned in any existing
  German lexica and did not comply with inflectional rules of German
  grammar.\label{fw}
\end{enumerate}
Though this division is admittedly arbitrary to a certain degree and
also has the disadvantage of simultaneously involving different
linguistic criteria, the underlying notion here was simple -- valid
words could have been omitted from an MRD either due to the
limitations of developers' capacities (group \ref{regular}), active
word formation processes or lexical productivity of the language
itself (groups \ref{compound} through \ref{new}) or also due to
language's openness to foreign language systems (groups \ref{loan} and
\ref{fw}).

@@

%% As is clear from the table, deletions are by far the most common type
%% of misspellings. The reasons for that are either relatively frequent
%% omissions of characters made by users or even more often automatic
%% truncations of too long messages which are performed by Twitter
%% service itself.\footnote{As is generally known, Twitter imposes a
%%   strong restriction on the length of posted messages which can be no
%%   longer than 140 characters. Upon exceeding this length, tweets get
%%   automatically truncated to the maximal allowed length.}

Finally, the third group -- ``Stylistic specifics of text genre'' --
was subdivided into the subgroups:
\begin{enumerate}
  \item @-tokens, e.g. \textit{@ZDFonline}, \textit{@sechsdreinuller};
  \item hashtags, e.g. \textit{\#Kleinanzeigen}, \textit{\#wetter};
  \item links, e.g. \textit{http://t.co}, \textit{sueddeutsche.de};
  \item smileys, e.g. \textit{:-P}, \textit{xD};
  \item slang, e.g. \textit{OMG}, \textit{WTF} etc.
\end{enumerate}
according to the formal or lexical class which the tokens belonged
to. Additionally, we also considered as \textit{slang} spelling
variants of standard language words which could be regarded as their
colloquial equivalents. Such words included cases like, for example,
\textit{ned} instead of \textit{nicht} or \textit{grad} instead of
\textit{gerade}, and were marked both as \textit{misspelling}s and
\textit{slang} in our classification. A detailed statistics on the
subgroups of class 3 is shown in Table \ref{table:style}:
\begin{table}
  \caption{Distribution of OOV words belonging to the class
    ``Stylistic specifics of text genre''\label{table:style}}
  \begin{tabular}{p{0.4\textwidth}*{4}{>{\centering\arraybackslash}p{0.15\textwidth}}}
    \hline\noalign{\smallskip}
    \multirow{2}{*}{OOV subclass} & %
    \multicolumn{2}{c}{\texttt{hunspell}} & %
    \multicolumn{2}{c}{\texttt{TreeTagger}}\\
    & \totaloov{} & \uniqoov{} & \totaloov{} & \uniqoov{}\\
    \noalign{\smallskip} \hline
    @-tokens & 13.02 & 20.23 & 16.19 & 21.91\\
    hashtags & 7.35 & 6.18 & 13.06 & 10.59\\
    links & 2.43 & 0.4 & 4.89 & 6.07\\
    smileys & 2 & 0.73 & 6.88 & 1.2\\
    slang & 16.22 & 5.27 & 6.94 & 4.77\\\hline
    {\bfseries total} & 41.02 & 32.81 & 47.96 & 44.54\\
    \noalign{\smallskip} \hline
  \end{tabular}
\end{table}

A striking outlier of 16.22~\% for slang tokens in column 1 of the
Table is explained by the fact that the word ``RT'' which occurred
1,235 times in our texts and was by far the most frequent OOV in
analyzed data set, was recognized as out-of-vocabulary by
\texttt{hunspell} but was not deemed as such by
\texttt{TreeTagger}. Luckily, such singleton cases were rather rare
exceptions during the analyses and did not affect much the remaining
classes of OOVs, so that distributions of unknown tokens for both
tools were more or less similar or at least comparable to a certain
extent.

However, an even more important conclusion which can be drawn from the
analyzed data is the fact that for both \texttt{TreeTagger} and
\texttt{hunspell}, Twitter-specific phenomena like special tokens,
colloquial expressions or sloppily typed words accounted for more than
a half of all unknown words found during the analysis. Since different
classes of these Twitter phenomena were formed by different processes
and also have different degrees of ambiguity, we will suggest
different normalization procedures for each of them in the next
Section.

@@

\subsection{Replacement of Twitter-Specific Phenomena}
According to Parker (\cite{parker}), hasthags were supposedly
introduced by Chris Messina, a renowned advocate of open-source
community, in August 2007. The ``hash godfather'', as Messina
describes himself, presumably borrowed the idea of marking relevant
topics with the ``\#''- sign from Internet Relay Chats (IRC) -- the
forebears of modern social networks -- which have been using the pound
character for marking their channel names since the early 1990s.

Luckily for us, this ``novelty'' by Messina brought along a strict
formal feature by which future hashtags could be identified. Other
types of words which are usually considered as OOVs in Twitter but
also have some unambiguous formal traits include at-tokens,
hyperlinks, e-mail addresses, and smileys. The presence of such formal
criteria suggests that these classes could best be handled by
rule-based methods and namely finite-state techniques like
finite-state transducers (FST) (cf. Jurafsky and Martin,
\cite{jurafsky}, pp.57-60).

For our purposes, we developed a prototypic Python system analogous to
an FST in which a set of regular expressions was associated with
corresponding actions performed on matched subgroups in cases when
these subgroups did not overlap. If some rules clashed together on
certain contexts, a set of special criteria was used to determine
which rule the preference should be given.

In this system, unambiguous tokens like web links and e-mail addresses
were replaced with an artificial word ``\%Mail''. Emoticons were
substituted by tokens ``\%PosSmiley'' or ``\%NegSmiley'', depending on
the type of emotion presumably conveyed by these expressions. In cases
when an emoticon was ambiguous with regard to its polarity, it was
replaced with the special word ``\%Smiley''. Furthermore, leading
``\#{}'' characters were stripped off from the beginning of hashtags
thus leaving only the alphabetic part of these tokens.

A more complicated case turned out to be the at-tokens and
hyperlinks. For them, we had to disambiguate whether these words
played a certain syntactic role in a sentence or could be easily
omitted without breaking the syntactic structure. In the former case,
these words were replaced with artificial counterparts. In the latter
case, such words were deleted from sentence in order not to confuse
further intended parsing processing.

@@

All the artificial words were added to custom user dictionaries of
\texttt{TreeTagger} and \texttt{hunspell}. For tagging, we assigned NE
tags to all artificial tokens with the only exception of emoticon
replacements which got the tag INJ. Furthermore, positions and lengths
of all replacements we remembered along with original words which were
replaced and a restoration step to original forms was performed for
certain classes of tokens after tagging.

%
\subsection{Restoration of misspellings}

%
\section{Evaluation}\label{evaluation}
